\documentclass[11pt]{article}
\usepackage{enumerate}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathabx}
\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{parskip}
\usepackage{commath}

\begin{document}

\title{Com S 331 Assignment 4}
\author{Adam Hammes $\bullet$ hammesa@iastate.edu }
\maketitle

\section*{Problem 33}

\begin{enumerate}[(a)]
	\item b*(ab*ab*)*
	\item a*ba*(ba*ba*)*
	\item b*(ab*ab*)* + a*ba*(ba*ba*)*

\end{enumerate}

\section*{Problem 34}

See Figures 34a, 34b, 34c

\section*{Problem 35}
Let $M = (Q,\ \Sigma,\ \delta,\ s,\ F)$ be a DFA for $A$.

General intuition of an NFA $M'$ that solves MiddleThirds $A$: We have five pebbles- red, blue, green, yellow, black- on the automaton at any time. The red pebble starts on $s$, the blue/green pebbles on a nondeterministically chosen state of $M$, and the yellow/black pebbles on another such chosen state. While the pairs of pebbles each start in the same place, the blue and yellow pebbles never change state; they are used as "markers" for the 1/3 and 2/3 parts of the string.

As we read in a string $x$, the red and yellow pebbles move forward according to some nondeterministically chosen symbol; the green pebble moves forward according to the input symbol. We accept a string if three conditions are met:
\begin{enumerate}
	\item The red pebble occupies the same state as the blue pebble
	\item The green pebble occupies the same state as the yellow pebble
	\item The black pebble is in an accepting state of $M$
\end{enumerate}

Since the green pebble follows the input symbols, it represents our string $x$  By the three conditions, we know that there are two strings $m$ and $n$, represented by the path that the red and black take, such that $|m| = |x| = |n|$. We also know from 1 and 2 that $mxn$ represents a string in $M$, thus the string $x$ which we accepted is in MiddleThirds $M$. Let's define this NFA, $M' = (Q',\ \Sigma,\ \Delta',\ S', F')$, more formally.

\begin{description}
	\item[$Q'$] We need to keep track of pebbles of (potentially) any arrangement of five pebbles; thus $Q' = Q^5$.
	\item[$S'$] First, our red pebble starts at the beginning, or $s$. There is no restriction of the blue/green pebbles or the black/yellow pebbles, other than they start at the same state, so $S' = \{(s,\ a,\ a,\ b,\ b) \mid a,b \in Q\}$.
	\item[$F'$] Our red pebble has caught up our blue pebble, and the same with green and yellow. Black is an accepting state. $F' = \{(a,\ a,\ b,\ b,\ c) \mid c \in F \}$.
	\item[$\Delta'$] Two pebbles, blue and yellow, don't change no matter the input. The green pebble, following $x$, follows the original $\delta$ for the input symbols. Finally, the red and black pebbles go to the set of states that *any* input symbol could lead them to.
	\[ \Delta'((u,\ v,\ w,\ y,\ z),\ a) = \{(\delta(u,b),\ v,\ \delta(w,a),\ y,\ \delta(z,c)) \mid b,c \in \Sigma \} \]

\end{description}

A lemma relating $\hat{\Delta}$ to $\hat{\delta}$ will be useful in our final proof, so I prove one next.



Lemma 1: For any $x \in \Sigma^*$, 
	\[ \hat{\Delta}'(S', x) = \{(\hat{\delta}(s,m),\ a,\ \hat{\delta}(a,x),\ b,\ \hat{\delta}(b,n)\ \mid m,n \in \Sigma^*\}\]
	
	Proof by induction on $x$. Base case: 
\begin{align*}
	\hat{\Delta}'(S', \lambda) &= S' \\
	&= \{(s,a,a,b,b) \mid a,b \in Q\} \\
	&= \{ ( \hat{\delta}(s,m),\  a,\ \hat{\delta}(a, \lambda),\ b,\ \hat{\delta}(b,n)) \mid m,n  \in \Sigma^0\}\ \text{\checkmark}	
\end{align*}

Inductive step: Assume
	\[\hat{\Delta}'(s',\ x) = \{ (\hat{\delta}(s,m),\ a,\ \hat{\delta}(a,x),\ b,\ \hat{\delta}(b,n)) \mid a,b \in Q \wedge m,n \in \Sigma^{|x|}\} \]
	
Prove
	\[\hat{\Delta}'(s',\ xc) = \{ (\hat{\delta}(s,md),\ a,\ \hat{\delta}(a,xc),\ b,\ \hat{\delta}(b,ne)) \mid a,b \in Q \wedge m,n \in \Sigma^{|x|}\} \]
	
$\hat{\Delta}(S', xc)$

\begin{align*}
&= \bigcup _{(u,a,w,b,y) \in \hat{\Delta}(S', x)} \hat{\Delta}'(\hat{\Delta}'(x),\ c)
\\
&= \bigcup _{a,b\ \in Q,\ n,m\ \in\ \Sigma^{|x|}} \Delta'((\hat{\delta}(s,m),\ a,\ \delta(\hat{\delta}(a, x),c),\ b, \delta(\hat{\delta}(b,n)), c) 
\\
&= \bigcup _{a,b\ \in Q,\ n,m\ \in\ \Sigma^{|x|}} \{ \delta(\hat{\delta}(s,m),d),\ a,\ \delta(\hat{\delta}(a, x), c),\ b,\ \delta(\hat{\delta}(b,n),e) \mid d,e \in \Sigma \}
\\
&= \{ ( \hat{\delta}(s, md),\ a,\ \hat{\delta}(a, xc),\ b,\ \hat{\delta}(b, ne)) \mid a,b \in Q \wedge m,n \in \Sigma^{|x|} \wedge d,e \in \Sigma \} \text{\checkmark}
\end{align*}

We now can prove that L(M') = MiddleThirds L(M). For any $x \in \Sigma^*$, $x\in $ L(M')
\begin{align*}
&\iff \hat{\Delta}'(S', x)\ \bigcap \ F' \neq \emptyset
\\
&\iff \{( \hat{\delta}(s,m), a, \hat{\delta}(a, x),\ b,\ \hat{\delta}(b,n) ) \mid n,m \in \Sigma^{|x|},\ a,b \in Q \}
\\
&\bigcap\ \{w,\ w,\ u,\ v,\ v) \mid w,v \in Q\ u \in F' \} \neq \emptyset
\\
&\iff \exists m,n \in \Sigma^{|x|},\ a,b \in Q\ \hat{\delta}(s,m) = a,\ \hat{\delta}(a, x) = b,\ \hat{\delta}(b, n) \in F
\\
&\iff \exists m,n \in \Sigma^{|x|}\ \hat{\delta}(\hat{\delta}(\hat{\delta}(s, m), x), n) \in F
\\
& \iff \exists m,n \in \Sigma^{|x|}\ \hat{\delta}(s, mxn) \in F
\\
& \iff \exists m,n \in \Sigma^{|x|}\ mxn \in \text{ L(M)}
\\
&\iff x \in \text{ MiddleThirds(M) \checkmark}
\end{align*}

\section*{Problem 36}

See Figures 36a, 36b

\section*{Problem 37}

Note: I am sorry that I couldn't figure out how to simplify my regular expressions any further- I know how hard this must be to grade.

a(b+aa)*+(ab*a)*b(a+b(ab*a)*b)*b(ab+a)a +(b(ab*a)*a)*

\section*{Problem 38}

a(aa+bb)*+(a(aa+bb)*ab+b)(b(a(b(b+a))*a)*b +a(b(aa+ab)*)b)*a)*+(ba+ab)(aa+bb)*

\section*{Problem 39}

\begin{enumerate}[(a)]
	\item b*
	\item b*a*
	
\end{enumerate}

\section*{Problem 40}

\begin{enumerate}[(a)]
	\item (i)
	\item (ii)
	\item (v)
	\item (iv)
	\item (iii)
	
\end{enumerate}

\end{document}