\documentclass[11pt]{article}

\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{parskip}

\let\iff\leftrightarrow
\let\imp\rightarrow

\begin{document}

\title{Com S 331 Homework 12}
\author{Adam Hammes $\bullet$ hammesa@iastate.edu}
\date{December 9, 2014}
\maketitle

Note: $REG$ stands for the set of all regular languages, $DEC$ all decidable languages, $CE$ computable, and $co-CE$ co-computable.


\section*{Problem 89}

Let \textbf{rev} $L \subseteq \Sigma^*$ be as defined on page 302 of the text; that is,
	\[\text{\textbf{rev }} L = \{\text{\textbf{rev }} \mid x \in L\}\]
where \textbf{rev} $x \in \Sigma^*$ is the reverse of the string. Let $B$ = \textbf{rev} $A$.
I prove $B$ is not regular.

Let $c \in \mathbb{N}$.
Choose $n \in \mathbb{Z}^+$ such that $C(0^n) > c$, and let $x = 0^n1$.
The biggest possible value for $k$ in the string is $n$; thus we need at least another $n$ zeroes to get a valid string.
Thus $y^{B} _{x,1} = 0^n10^n$ and 
	\begin{align*}
		C( y^{B} _{x,1} ) &= C( 0^n) \\
		&> c + \log 1\ \checkmark
	\end{align*}

We proved on problem 18 of the homework that if a set $L$ is regular then \textbf{rev} $L$ is also regular.
Since $B$ is not regular, $A$ is not regular by contrapositive. $\checkmark$


\section*{Problem 90}

Consider $s \in A$.
By definition of $A$, $s$ is of the form $0^kx$ where $\#(0,x) \geq k$ and $k>0$.
Note that $x$ must contain at least one zero.

Let $y$ be $0^{k-1}x$. $x$ has at least one 0, so $0y \in A$.
Also, $0y = s$.
Because $k >0$, such a $y$ exists for every $s \in A$.
Thus we can can express the language $A$ as follows:
	\[ A = \{0x \mid \#(0,x) > 0 \}\]
which is decided by the regular expression
	\[ 0\ (0+1)^*\ 0\ (0+1)^*\]


\section*{Problem 92}

Let $P_\infty$ be the following property:
	\[ P_\infty = \{ n \mid L(M_n) \text{ is infinite}. \}\] 

Let $M_A$ and $M_B$ be Turing mahcines, with $M_A$ accepting on all inputs and $M_B$ rejecting on all inputs.
Then L($M_A$) is finite and L($M_B$) is infinite. 
Thus $P_\infty$ is non-trivial.
Since $P_\infty$ is non-trivial, by Rice's theorem $INF$ is undecidable. $\checkmark$



\section*{Problem 93}

Suppose for contradiction that $T$ is computable.
Let $M_K$ be as follows:
\begin{itemize}
	\item input $n \in \mathbb{N}$
	\item let $a = T(n)$
	\item run $M_n(n)$; if in that time $M_n(n)$ halts, ACCEPT else REJECT
\end{itemize}

Because $T(n)$ places an upper limit on how long a halting Turing machine will take to halt, we can definitely tell if $M_n(n)$ halts.
Thus $M_K$ decides the diagonal halting problem, a contradiction. $\checkmark$


\section*{Problem 94}

Consider $R^C$,
	\[ \{x \in \{0,1\}^* \mid C(x) < |x| \}. \]
I prove that $R^C \in CE$, equivalent to $R \in co-CE$.

Let $U$ be a universal Turing machine.
Define $M_R$ as follows:
\begin{itemize}
	\item input $x \in \{0,1\}^* $
	\item let $A$ be $\{0,1\}^{\leq |x|}$
	\item for $i \in \mathbb{Z}^+$ from 1 to $\infty$:
	\begin{itemize}
		\item for $a \in A$, run $U(a)$ for $i$ steps
		\item if $U$ outputs $x$, ACCEPT
	\end{itemize}
\end{itemize}

Consider $x \in R^C$.
Because $C(x) < |x|$, there exists a Turing machine $\pi$ such that $|\pi| < x$ and $U(\pi) = x$.
$M_R$ iterates through every possible encoding of a Turing machine with length less than $|x|$, accepting if and only if the universal Turing machine outputs $x$ on that encoding.
Therefore $L(M_R) = R^C$. $\checkmark$



\section*{Problem 96}

From 95 we know that every c.e. subset of $R$ is finite.
Because $R$ is infinite and a subset of itself, we know that $R \notin CE$.
Since $DEC \subseteq CE$, we can conclude that $R \notin DEC$. $\checkmark$.


\end{document}